---
title: "Covid91 vaccine study Final2023"
author: "Daniela Olvera-Mota"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ The Covid 921 vaccine does not exhibit a significant difference in effectiveness between male populations.

$H_A$ The Covid 921 vaccine does exhibit a significant difference in effectiveness between male populations.

```{r}
library(dplyr)
Males = filter (FinalData, FinalData$sex == 'M')
```

### Methods

Males is the categorical variable , Covid 921 vaccine administration is an explanatory variable.Covid 921 preventative effectiveness is a response variable.

Therefore, we will use the cat_cat helper files.

### Results 

### Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

According to the bar graph the males who got the placebo had a higher percentage catching covid 19.Opposed to the males who received the vaccine 921 which had a lower percentage of catching covid.Among the males who got covid ,the people who received the vaccine 921 had significant reduction in catching covid.Compared to the males who received the placebo, who had a higher percentage of catching covid.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

### Discussion and Conclusion


## Females

$H_0$ The Covid 921 vaccine does not exhibit a significant difference in effectiveness between female populations.

$H_A$ The Covid 921 vaccine does exhibit a significant difference in effectiveness between female populations.

```{r}
library(dplyr)
Females = filter (FinalData, FinalData$sex == 'F')
```

### Methods

Females is the categorical variable , Covid 921 vaccine administration is an explanatory variable.Covid 921 preventative effectiveness is a response variable.

Therefore, we will use the cat_cat helper files.

### Results 

### Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

### Numerical Descriptive Results

### Inferential Results

### Discussion and Conclusion

## LGBTQ

$H_0$ The Covid 921 vaccine does not exhibit a significant difference in effectiveness between LGBTQ populations.

$H_A$ The Covid 921 vaccine does exhibit a significant difference in effectiveness between LGBTQ populations.

```{r}
library(dplyr)
LGBTQ = filter (FinalData, FinalData$LGBTQ == 'gay')
```

### Methods

LGBTQ is the categorical variable , Covid 921 vaccine administration is an explanatory variable.Covid 921 preventative effectiveness is a response variable.

Therefore, we will use the cat_cat helper files.

### Results 

### Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

### Numerical Descriptive Results

### Inferential Results

### Discussion and Conclusion

## Druggies

$H_0$ The Covid 921 vaccine does not exhibit a significant difference in effectiveness between druggies populations.

$H_A$ The Covid 921 vaccine does exhibit a significant difference in effectiveness between druggies populations.

```{r}
library(dplyr)
Druggies = filter (FinalData, FinalData$DrugUser == 'yes')
```

### Methods

Druggies is the categorical variable , Covid 921 vaccine administration is an explanatory variable.Covid 921 preventative effectiveness is a response variable.

Therefore, we will use the cat_cat helper files.

### Results 

### Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment ,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")
```

### Numerical Descriptive Results

### Inferential Results

### Discussion and Conclusion


# Overall Results and Conclusions